```{r}
#| include: false
source("_setup.R")
```

# Introduction

## Background

Epitranscriptomics encompasses all post-transcriptional biochemical modifications, with over 160 different types of modifications identified. Among these, N6-methyladenosine (m6A) is a chemical derivative of the adenosine molecule where the 6th nitrogen is methylated. This m6A RNA methylation is prevalent in most eukaryotes and has been detected in numerous species. Furthermore, m6A is recognized as one of the most abundant mRNA modifications in human transcripts.

Since its initial discovery, numerous new findings have shed light on the roles and functions of m6A. Notably, m6A is well recognized for its involvement in post-transcriptional gene expression regulation, a crucial biological process in all living organisms. It exerts various regulatory mechanisms by affecting multiple aspects of RNA metabolism, including alternative RNA splicing, alternative polyadenylation, RNA nuclear export, and mRNA degradation, stabilization, and translation. The regulation of gene expression mediated by m6A has significant implications in numerous physiological and pathophysiological processes, such as brain development, cardiac homeostasis, immune system development, spermatogenesis, and skeletal function. Moreover, recent studies have provided extensive evidence of m6A's role in various cancers and its influence on immune activity related to tumor micro-environment.

The basic mechanism of m6A can be explained through the involvement of three key players: the writer, the eraser, and the reader. The writer, also known as methyltransferase, adds a methyl group to the normal adenosine (A) molecule, converting it into m6A. Conversely, the eraser, which acts as a demethylase, removes the methyl group from m6A, converting it back to A. The reader refers to specific RNA-binding proteins (RBPs) which recognize and bind to m6A. Upstream regulation involves the functions of writer and eraser, addressing why and how methylation and demethylation processes occur. On the other hand, downstream regulation refers to the diverse functions induced by different readers binding to m6A.

m6A not only regulates various normal cellular functions, including cell development, but also plays a significant role in diseases such as cancer and viral infections. The condition-specific m6A epitranscriptome is crucial in many diseases, particularly in cancers. Extensive efforts have been made to profile the m6A epitranscriptome and elucidate the molecular mechanisms involving m6A regulators, known as Writers, Erasers, and Readers (WERs). Dysregulation of m6A WERs has been observed in more than 17 different types of cancer. Consequently, there has been a surge in research focus on m6A, resulting in over 2000 publications related to m6A and 1000 publications specifically addressing the association between m6A and cancer in 2022, according to PubMed record.

Elucidating the determinants of m6A deposition at specific loci in the genome is critical to understanding condition-specific and conserved m6A regulation of gene expression. Several studies have addressed context-dependent and context-independent m6A deposition and their regulatory mechanisms. A recent review paper [@TC.m6A.review] explained the regulation of m6A deposition on mRNA and suggests its context-dependent and context-independent effects on mRNA, particularly concerning mRNA decay and translation. The presence of wide-spread m6A differential methylation in various mammalian cellular conditions implies the possible involvement of extrinsic m6A factors, such as RBPs or mRNA secondary structures. Numerous studies have demonstrated context-specific m6A regulation under different conditions. However, the exploration of context-independent (TC) m6A mRNA modification has not been thoroughly conducted. A recent study [@m6A.mazter.seq] reveals that m6A in the yeast genome is primarily determined TCally and "hard-coded" by TC cis-sequences from MAZTER-seq data. However, the extent to which TC and context-dependent regulators deposit m6A in mammalian cells has yet to be investigated.


## Aims

```{r}
sample.list = readRDS(paste0(cloud, "/data/sample.meta/sample.list.after.drop.rds")) %>% subset(., source != "gtex") %>% subset(., category=="tissue")
N.all.samples = nrow(sample.list)
N.all.tissues = length(unique(sample.list$class))
```

In this study, we examined `r max(N.all.samples)` human tissue methylated RNA immunoprecipitation sequencing (MeRIP-seq) data derived from `r max(N.all.tissues)` tissues to investigate the presence of TC m6A sites. TC m6A sites are deposited under normal physiological conditions and play important roles in regulating normal cellular functions. We identified TC m6A sites associated with context-independent m6A gene expression regulation, characterized their distinct features, and explored their regulatory mechanisms.

Following is the aims of the study:

1.  Investigation of context independent (TC) m6A deposition and its regulatory mechanism related with m6A mRNA modification
- Identification of TC m6A sites
- Factors define deposition of TC m6A sites
- Upstream regulation of TC m6A sites causing m6A modification
- Downstream function of TC m6A sites affecting on mRNA

2.  Define TC/extrinsic m6A sites

3.  Find genetic features/marks of TC m6A sites
- Methylation levels
- Gene expression levels
- Compare with m6A atlas sites
- Compare with Whistle sites
- Neighboring hood motif search
- comparison between TC and infrequent sites

4.  Study the function of TC m6A sites
- Why the TC sites are there?
- How are these TC sites methylated?
- What is the function of these TC sites?

5.  Is there a difference in their sequences between TC and non-TC sites?
- a set of unique sequence motifs that are bound by unique RPBs
- train a deep learning model to classify TC and non-TC site sequences

6.  cis-regulating cofactors associated with different readers (e.g. YTHDF vs IGF2BP)
- all readers recognize DRACH motif but have different functions
- YTHDF promotes mRNA decay but IGF2BP promotes stability
- use DNABert to perform a classification between sequences bound by YTHDF and IGF2BP


